---
title: "A Short Tour"
subtitle: "of the Predictive Modeling Process"
author: "Bo Werth"
date: "2017/02/07"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

## Case Study: Predicting Fuel Economy

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r tour_startup,echo=FALSE,message=FALSE,warning=FALSE}
library(knitr)
opts_chunk$set(digits = 3, tidy = FALSE, fig.path = '02-A-Short-Tour/')
library(AppliedPredictiveModeling)
library(caret)
library(earth)
library(lattice)
```

---

## Format data for plotting against engine displacement

- Sort by engine displacement
- Combine data into one data frame

```{r fuel_format, echo=TRUE}
data(FuelEconomy)
cars2010 <- cars2010[order(cars2010$EngDispl),]
cars2011 <- cars2011[order(cars2011$EngDispl),]
cars2010a <- cars2010
cars2010a$Year <- "2010 Model Year"
cars2011a <- cars2011
cars2011a$Year <- "2011 Model Year"
plotData <- rbind(cars2010a, cars2011a)
```

---

## `xyplot`

```{r fuel_plot, fig.width=9.5, fig.height=5.5, tidy=FALSE, message=FALSE, warning=FALSE}
xyplot(FE ~ EngDispl|Year, plotData,
       xlab = "Engine Displacement",
       ylab = "Fuel Efficiency (MPG)",
       between = list(x = 1.2))
```

---

## Single Linear Model

Fit a single linear model and conduct 10-fold CV to estimate the error

```{r linear, tidy=TRUE, message=FALSE, warning=FALSE}
set.seed(1)
lm1Fit <- train(FE ~ EngDispl, 
                data = cars2010,
                method = "lm", 
                trControl = trainControl(method= "cv"))
lm1Fit
```

---

## Fit a quadratic model too

```{r square, tidy=TRUE, message=FALSE, warning=FALSE}
cars2010$ED2 <- cars2010$EngDispl^2 # Create squared terms
cars2011$ED2 <- cars2011$EngDispl^2
set.seed(1)
lm2Fit <- train(FE ~ EngDispl + ED2, 
                data = cars2010,
                method = "lm", 
                trControl = trainControl(method= "cv"))
lm2Fit
```

---

## Finally a MARS model (via the earth package)

```{r mars, tidy=TRUE, message=FALSE, warning=FALSE}
set.seed(1)
marsFit <- train(FE ~ EngDispl, 
                 data = cars2010,
                 method = "earth",
                 tuneLength = 15,
                 trControl = trainControl(method= "cv"))
marsFit

plot(marsFit)
```

---

# Predict and Performance

## Predict the test set data

```{r predict, tidy=FALSE, message=FALSE, warning=FALSE}
cars2011$lm1  <- predict(lm1Fit,  cars2011)
cars2011$lm2  <- predict(lm2Fit,  cars2011)
cars2011$mars <- predict(marsFit, cars2011)
```

---

## `postResample`

Get test set performance values via caret's `postResample` function

```{r performance, tidy=FALSE, message=FALSE, warning=FALSE}
postResample(pred = cars2011$lm1,  obs = cars2011$FE)
postResample(pred = cars2011$lm2,  obs = cars2011$FE)
postResample(pred = cars2011$mars, obs = cars2011$FE)
```
