---
title: "Chapter 3"
subtitle: "Data Pre-Processing"
author: "Bo Werth"
date: "2017/02/07"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

## Case Study: Cell Segmentation in High-Content Screening

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r tour_startup,echo=FALSE,message=FALSE,warning=FALSE}
library(knitr)
opts_chunk$set(digits = 3, tidy = FALSE, fig.path = '03-Data-Pre-Processing/')
library(AppliedPredictiveModeling)
library(caret)
library(e1071)
library(ellipse)
library(corrplot)
```

---

## Cell Segmentation Data

```{r cell_data}

data(segmentationOriginal)

## Retain the original training set
segTrain <- subset(segmentationOriginal, Case == "Train")

## Remove the first three columns (identifier columns)
segTrainX <- segTrain[, -(1:3)]
segTrainClass <- segTrain$Class
```

---

## Data Transformations for Individual Predictors

The column VarIntenCh3 measures the standard deviation of the intensity of the
pixels in the actin filaments

```{r cell_transform}
max(segTrainX$VarIntenCh3)/min(segTrainX$VarIntenCh3)
skewness(segTrainX$VarIntenCh3)
```

---

## `preProcess`

Use caret's preProcess function to transform for skewness

```{r cell_skewness}
segPP <- preProcess(segTrainX, method = "BoxCox")
## Apply the transformations
segTrainTrans <- predict(segPP, segTrainX)
```

---

## VarIntenCh3

Results for a single predictor

```{r inten}
segPP$bc$VarIntenCh3
```

---

## VarIntenCh3 Natural Units

```{r inten_hist_nat, fig.height=5.5}
histogram(~segTrainX$VarIntenCh3,
          xlab = "Natural Units",
          type = "count")
```

---

## VarIntenCh3 Log Units

```{r inten_hist_log, fig.height=5.5}
histogram(~log(segTrainX$VarIntenCh3),
          xlab = "Log Units",
          ylab = " ",
          type = "count")
```

---

## PerimCh1

```{r perim}
segPP$bc$PerimCh1
```

---

## PerimCh1 Natural Units

```{r perim_hist_nat, fig.height=5.5}
histogram(~segTrainX$PerimCh1,
          xlab = "Natural Units",
          type = "count")
```

---

## PerimCh1 Transformed Data

```{r perim_hist_log, fig.height=5.5}
histogram(~segTrainTrans$PerimCh1,
          xlab = "Transformed Data",
          ylab = " ",
          type = "count")
```

```{r, echo=FALSE, eval=FALSE}
################################################################################
### Section 3.3 Data Transformations for Multiple Predictors

## R's prcomp is used to conduct PCA
pr <- prcomp(~ AvgIntenCh1 + EntropyIntenCh1,
             data = segTrainTrans,
             scale. = TRUE)

transparentTheme(pchSize = .7, trans = .3)

xyplot(AvgIntenCh1 ~ EntropyIntenCh1,
       data = segTrainTrans,
       groups = segTrain$Class,
       xlab = "Channel 1 Fiber Width",
       ylab = "Intensity Entropy Channel 1",
       auto.key = list(columns = 2),
       type = c("p", "g"),
       main = "Original Data",
       aspect = 1)

xyplot(PC2 ~ PC1,
       data = as.data.frame(pr$x),
       groups = segTrain$Class,
       xlab = "Principal Component #1",
       ylab = "Principal Component #2",
       main = "Transformed",
       xlim = extendrange(pr$x),
       ylim = extendrange(pr$x),
       type = c("p", "g"),
       aspect = 1)


## Apply PCA to the entire set of predictors.

## There are a few predictors with only a single value, so we remove these first
## (since PCA uses variances, which would be zero)

isZV <- apply(segTrainX, 2, function(x) length(unique(x)) == 1)
segTrainX <- segTrainX[, !isZV]

segPP <- preProcess(segTrainX, c("BoxCox", "center", "scale"))
segTrainTrans <- predict(segPP, segTrainX)

segPCA <- prcomp(segTrainTrans, center = TRUE, scale. = TRUE)

## Plot a scatterplot matrix of the first three components
transparentTheme(pchSize = .8, trans = .3)

panelRange <- extendrange(segPCA$x[, 1:3])
splom(as.data.frame(segPCA$x[, 1:3]),
      groups = segTrainClass,
      type = c("p", "g"),
      as.table = TRUE,
      auto.key = list(columns = 2),
      prepanel.limits = function(x) panelRange)

## Format the rotation values for plotting
segRot <- as.data.frame(segPCA$rotation[, 1:3])

## Derive the channel variable
vars <- rownames(segPCA$rotation)
channel <- rep(NA, length(vars))
channel[grepl("Ch1$", vars)] <- "Channel 1"
channel[grepl("Ch2$", vars)] <- "Channel 2"
channel[grepl("Ch3$", vars)] <- "Channel 3"
channel[grepl("Ch4$", vars)] <- "Channel 4"

segRot$Channel <- channel
segRot <- segRot[complete.cases(segRot),]
segRot$Channel <- factor(as.character(segRot$Channel))

## Plot a scatterplot matrix of the first three rotation variables

transparentTheme(pchSize = .8, trans = .7)
panelRange <- extendrange(segRot[, 1:3])

upperp <- function(...)
  {
    args <- list(...)
    circ1 <- ellipse(diag(rep(1, 2)), t = .1)
    panel.xyplot(circ1[,1], circ1[,2],
                 type = "l",
                 lty = trellis.par.get("reference.line")$lty,
                 col = trellis.par.get("reference.line")$col,
                 lwd = trellis.par.get("reference.line")$lwd)
    circ2 <- ellipse(diag(rep(1, 2)), t = .2)
    panel.xyplot(circ2[,1], circ2[,2],
                 type = "l",
                 lty = trellis.par.get("reference.line")$lty,
                 col = trellis.par.get("reference.line")$col,
                 lwd = trellis.par.get("reference.line")$lwd)
    circ3 <- ellipse(diag(rep(1, 2)), t = .3)
    panel.xyplot(circ3[,1], circ3[,2],
                 type = "l",
                 lty = trellis.par.get("reference.line")$lty,
                 col = trellis.par.get("reference.line")$col,
                 lwd = trellis.par.get("reference.line")$lwd)
    panel.xyplot(args$x, args$y, groups = args$groups, subscripts = args$subscripts)
  }
splom(~segRot[, 1:3],
      groups = segRot$Channel,
      lower.panel = function(...){}, upper.panel = upperp,
      prepanel.limits = function(x) panelRange,
      auto.key = list(columns = 2))

################################################################################
### Section 3.5 Removing Variables

## To filter on correlations, we first get the correlation matrix for the 
## predictor set

segCorr <- cor(segTrainTrans)

corrplot(segCorr, order = "hclust", tl.cex = .35)

## caret's findCorrelation function is used to identify columns to remove.
highCorr <- findCorrelation(segCorr, .75)

################################################################################
### Section 3.8 Computing (Creating Dummy Variables)

data(cars)
type <- c("convertible", "coupe", "hatchback", "sedan", "wagon")
cars$Type <- factor(apply(cars[, 14:18], 1, function(x) type[which(x == 1)]))

carSubset <- cars[sample(1:nrow(cars), 20), c(1, 2, 19)]

head(carSubset)
levels(carSubset$Type)

simpleMod <- dummyVars(~Mileage + Type,
                       data = carSubset,
                       ## Remove the variable name from the
                       ## column name
                       levelsOnly = TRUE)
simpleMod

withInteraction <- dummyVars(~Mileage + Type + Mileage:Type,
                             data = carSubset,
                             levelsOnly = TRUE)
withInteraction
predict(withInteraction, head(carSubset))
```
